I would like to thank everyone who made this thesis a successful endeavour.\\
My supervisor Ines Thiele for giving me the opportunity to join the MSP lab and work on challenging projects on the boundaries of the knowledge in the field, and for the mentorship and support during my Ph.D. All the past and present members of the MSP lab for being wonderful colleagues, in particular, Eugen, Berto, and Fede. My CET committee Prof. Balling and Prof. Krueger for their guidance and advice, and who continue to inspire me in terms of the scientific method and rigor. Prof. Balling cannot be praised enough for supporting international students and for making LCSB a major pole for biomedical research in Europe and the World. LCSB was an intellectually stimulating place for me to grow and learn. Prof. Serge Haan for continuously encouraging student initiatives. Thanks to his support, the development of student activities were only bounded by our imagination. The external committee Prof. Moerland and Dr. Kuepfer for accepting to review my work.\\
My friends and colleagues at the LCSB starting with Al-Jamiaa Al-Arabiya: Amer, Remon, and Siham for bringing bits of home to Luxembourg and for the moments of joy and laughter that we had together. Les loulous, these beautiful humans, that, to be honest, I do not even know how to emphasize how life-affirming was meeting them: Zoe, Aurélien, Simone, Edward, Watson, Fran\c{c}ois, Geoffrey, Maxime, Marie, Alice, Christophe, Justine, Debby, Nico, and Nicole. Thanks for the memorable moments we shared together. Dr. Johan Thunberg at the LCSB for saying 'yes, and' to the craziest of ideas. His positive attitude and mathematical rigor inspire me a lot. And most importantly the RSG Luxembourg student group that I co-founded and I am happy to see thriving and developing. I would like to thank Aishwarya for believing in this project as a way of giving back to the bioinformatics community in Luxembourg.\\
To all of my colleagues, friends, and mentors at the LCSB and the UL, thank you.